---
title: "KNN (K_Nearest_Neighbour)"
author: "Clifford Mwenda"
date: "2024-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



DATA PREPROCESSING

Importing Data Set
```{r }
setwd("C:/Users/Latitude/Desktop/Mushroom Classification")
dataset = read.csv("mushroom.csv", na.strings = c("", " ", "NA"))
str(dataset)
```

Encoding Categorical Variables in the Data_set
```{r}
library(dplyr)
dataset <- dataset %>%
  mutate(across(-c(cap.diameter, stem.height, stem.width), as.factor))
str(dataset)
```

Check NAs and drop columns with NAs
```{r}
n_nas = colSums(is.na(dataset))
(n_nas/61069) * 100

dataset = dataset %>% select(where(~ !any(is.na(.))))
str(dataset)
```

Rearrange Columns
```{r}
dataset = dataset[, c(2:12, 1)]
str(dataset)
```

Convert Categorical Variables to Numeric for KNN
```{r}
dataset$class = as.numeric(as.integer(dataset$class))
dataset$cap.shape = as.numeric(as.integer(dataset$cap.shape))
dataset$cap.color = as.numeric(as.integer(dataset$cap.color))
dataset$does.bruise.or.bleed = as.numeric(as.integer(dataset$does.bruise.or.bleed))
dataset$gill.color = as.numeric(as.integer(dataset$gill.color))
dataset$stem.color = as.numeric(as.integer(dataset$stem.color))
dataset$has.ring = as.numeric(as.integer(dataset$has.ring))
dataset$habitat = as.numeric(as.integer(dataset$habitat))
dataset$season = as.numeric(as.integer(dataset$season))

str(dataset)
```

Split Data Set into Training and Testing Data Sets
```{r}
library("caTools")
set.seed(123)
split = sample.split(dataset$class, SplitRatio = 0.75)
training_set = subset(dataset, split == TRUE)
test_set = subset(dataset, split == FALSE)
```

Feature Scaling
```{r}
training_set = scale(training_set) #Z Score
test_set = scale(test_set)

head(training_set)
```






FITTING AND EVALUATING THE MODEL

Fitting K-NN to the Training Set and Predicting the test set results
```{r}
library(class)
y_pred = knn(train = training_set[, -12], #training data without dependent variable
             test = test_set[, -12],#test data without dependent variable
             cl = training_set[, 12],#dependent variable of the training set
             k = 5)#number of neighbors
summary(y_pred)
```

Evaluate Predictions using prediction matrix
```{r}
cm = table(test_set[, 12],#vector of real values
           y_pred)#vector of predicted values
cm

accuracy = (6746 + 8410)/15267
accuracy #99.27%
```

Apply K-Fold Cross Validation
```{r}
library(Rfast)
mod <- knn.cv(nfolds = 10, y = training_set[, 12], x = training_set[, -12], k = 5, dist.type = "euclidean", type = "C", method = "average") 
mod#98.98%
```


